Typed Functional Genetic Programming

### 2013-05-06 tex/thesis.tex
[tex] Inhabitation tree and more ..

\section{Type}

@@

@@

	In other words context is a set of statements with distinct variables or constants as subjects.

@@

	\section{Term generating grammar}
	Inference rules are good for deriving statements of a form $\Gamma \vdash M : \sigma$, but our
	goal is slightly different; we would like to generate many $\lambda$-terms M for a given type 
	$\sigma$ and context $\Gamma$.

@@

Inference rule 2: 
\[
	\frac{ (f , \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha ) \in \Gamma \qquad
	       \Gamma \vdash M_1 : \tau_1 \quad
	       \dotsm \quad
	       \Gamma \vdash M_n : \tau_n        
	      }
	     {\Gamma \vdash (f M_1 \dots M_n) : \alpha}
\]
\\
... there is a grammar rule:
\[ 
	( \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large f }}
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \tau_1 , \Gamma )
	  \mbox{ \Vtextvisiblespace[1em] } 
	  \dots
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \tau_n , \Gamma )
	  \quad \bigg)
\]
\\

@@

Now we will introduce \textit{Inhabitation tree}, structure slightly different from
\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
We can think about Inhabitation tree as about unfolded Inhabitation machine.
\\\\
\textbf{TODO}: TALK ABOUT And-Or-graph \\
\textbf{TODO}: TALK ABOUT "Barendregt-like" subsection \ref{barlike}
\\\\
Let's see some example:

@@

@@

		\subsection{Inhabitation Machine}
	\section{Roadmap}
	\section{Conversion to SKI combinators}

>>>>

	\textbf{TODO} 
	\begin{itemize} 
		\item TALK ABOUT "parenthesis" conventions. 
	\end{itemize}

\section{Type}
\label{deftype}

@@

\textbf{TODO} 
\begin{itemize}
	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
	\item TALK ABOUT arrow/parenthesis conventions. 
\end{itemize}

@@

	In other words context is a set of statements with distinct variables or constants as subjects.
	\\\\
	\textbf{TODO:} TALK ABOUT Context represents library/building blocks.

@@

\section{Term generating grammar}

Inference rules are good for deriving statements of a form $\Gamma \vdash M : \sigma$, but our
goal is slightly different; we would like to generate many $\lambda$-terms M for a given type 
$\sigma$ and context $\Gamma$.

@@

Inference rule 2: 
\[
	\frac{ (f , \rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha ) \in \Gamma \qquad
	       \Gamma \vdash M_1 : \rho_1 \quad
	       \dotsm \quad
	       \Gamma \vdash M_m : \rho_m        
	      }
	     {\Gamma \vdash (f M_1 \dots M_m) : \alpha}
\]
\\
... there is a grammar rule:
\[ 
	( \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large f }}
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_1 , \Gamma )
	  \mbox{ \Vtextvisiblespace[1em] } 
	  \dots
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_m , \Gamma )
	  \quad \bigg)
\]
\\\\
\textbf{TODO} 
\begin{itemize}
	\item SHOW correctness of those inference rules by composing them of 
		  $E^{\rightarrow}$, $I^{\rightarrow}$ and \textit{axiom}.
	\item SHOW more examples of inference rules transformed into grammar rules.
	\item DESCRIBE general algorithm for this transformation.
	\item TALK ABOUT $\tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$ 
	\item TALK ABOUT $\beta \eta$-normal form which is generated by this method.
\end{itemize}

@@

Now we will introduce \textit{Inhabitation tree}, structure slightly different from
\textit{Inhabitation machine}, which was introduced in \cite{barendregt10} by Henk Barendregt.
We can think about Inhabitation tree as about unfolded Inhabitation machine.
The motivation for using Inhabitation trees is belief that it will help us 
reason about generation of $\lambda$-terms 
of a given type $\sigma$ and with a given context $\Gamma$.

\subsection{Definition of Inhabitation tree}

\textit{Inhabitation tree} is a \textit{rooted tree}, possibly infinite. 
It has two types of nodes:

\begin{samepage}
\begin{itemize}
  \item Type nodes   
  			- containing type $\sigma \in \mathbb{T}$ 
  			- aka "OR-node" , Nonterminal-node. 
  \item Symbol nodes 
  			%- containing $ x^{*} \in \mathfrak P \left(V\right) -blobost pač ordered $ or $c \in C$
  			- containing "$\lambda$-head" (nonempty finite sequence of variable names) or constant
  			  name. 
  			- aka "AND-node" , Terminal-node.
\end{itemize}
\end{samepage}

We construct Inhabitation tree for given type $\sigma$ and context $\Gamma$.\\
We will define Inhabitation tree by describing its construction for a given $(\sigma,\Gamma)$.
Notice that it will closely follow the rules from \ref{barlike}:

\begin{itemize}
\item The root of Inhabitation tree for $(\sigma,\Gamma)$ is 
      \textit{type node} with $\sigma$ as type.
\item All \textit{type nodes} have as child nodes only \textit{symbol nodes}. 
\item And all \textit{symbol nodes} have as child nodes only \textit{type nodes}. 
\end{itemize}

Now we will resolve the child nodes of the root node.

There are two cases of $\sigma$ (recall \ref{deftype}): 
\begin{description}
	\item[Atomic type] 
		$\sigma = \alpha $ where $\alpha \in A$. 
	\item[Function type] 
		$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
		where $n \geq 1, \alpha \in A$.
\end{description}

First case \textbf{Atomic type} --- i.e., $\sigma = \alpha$ where $\alpha \in A$:\\
For every $(f,\rho_1 \rightarrow \dots \rightarrow \rho_m \rightarrow \alpha) \in \Gamma$
where $\alpha \in A$ there is a child \textit{symbol node} of the root containing constant name $f$.
This symbol node containing $f$ has $m$ child subtrees corresponding to Inhabitation trees for 
$(\rho_1,\Gamma),\dots,(\rho_n,\Gamma)$.

~

Compare this case with corresponding grammar rule:
\[ 
	( \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large f }}
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_1 , \Gamma )
	  \mbox{ \Vtextvisiblespace[1em] } 
	  \dots
	  \mbox{ \Vtextvisiblespace[1em] } 
	  ( \rho_m , \Gamma )
	  \quad \bigg)
\]

Second case \textbf{Function type} --- i.e., 
$\sigma = \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha$
where $n~\geq~1, \alpha~\in~A$:\\
For every  $i \in \{1,\dots,n\}$ we create new \textit{variable name} $x_i$ which is not yet included in context $\Gamma$ as variable or constant name.

There is one and only one child \textit{symbol node} of the root containing "$\lambda$-head" 
$\lambda x_1 \dots x_n$ which stands for sequence of variable names $(x_1,\dots,x_n)$.
This symbol node containing $\lambda x_1 \dots x_n$
has one and only one child subtree corresponding to Inhabitation trees for 
$(\alpha,\Gamma \cup \{ (x_1,\tau_1) , \dots , (x_n,\tau_n) \})$.

~

Compare this case with corresponding grammar rule:
\[ 
	( \tau_1 \rightarrow \dots \rightarrow \tau_n \rightarrow \alpha , \Gamma )  \longmapsto
	\bigg( \mbox{ {\Large 
	$\lambda x_1 \dots x_n .$ 
	}}( \alpha , \Gamma \cup \{ (x_1,\tau_1),\dots,(x_n,\tau_n) \} ) \quad \bigg)
\]
~\\
Let's consider following $(\sigma,\Gamma)$ as a simple example:

@@

This particular $(\sigma,\Gamma)$ results in the following tree:\\

@@

~

Second example features our well known example:

\begin{align*}
\sigma =& ~ (\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau \\ 
\Gamma =& \{ \}
\end{align*}

Which results in following tree:

\Tree
[.\text{ $(\sigma \rightarrow \tau) \rightarrow \sigma \rightarrow \tau $ } 
	[.\textbf{$\lambda$ f x }	
		[.\text{ $\tau$ }		
			[.\textbf{f} 
				[.\text{ $\delta$ }
					\textbf{x}					
				]
			]
		]
	]
]

\subsection{And-or tree and searching in Inhabitation tree}
Let us consider following definition of 
\textit{And-or tree}\footnote{
\textbf{TODO}: Mention that on WIKI there is more general definition, but 
for our purposes is this one sufficient.}:

\textit{And-or tree} is a rooted tree where each node is labeled as either \textit{and-node} 
or\footnote{xor} \textit{or-node}.

By \textit{solving} And-or tree $T$ we mean finding $T'$
subtree of $T$ such that it follows these conditions: 
\begin{itemize}
	\item The root of $T'$ is the root of $T$.
	\item Each \textit{and-node} in $T'$ has all the child nodes as in $T$.
	\item Each \textit{or-node}  in $T'$ has precisely one child 
	      node.\footnote{\textbf{TODO}: MENTION why precisely one 
	      and not at least one ..or CHANGE the def. }   
\end{itemize}
~
Let us now consider following labeling of Inhabitation tree:

\begin{itemize}
  \item \textbf{Type nodes}   are labeled as \textbf{or-nodes}.   
  \item \textbf{Symbol nodes} are labeled as \textbf{and-nodes}.
\end{itemize}

This labeling has following justification:

Selection of exactly one
child node in \textit{type node} corresponds to selection of exactly one
grammar rule in order to rewrite nonterminal symbol.

Selection of all the child nodes in \textit{symbol node} corresponds to 
rewriting all the nonterminal symbols in string that is being generated.

~\\\\
\textbf{TODO}: TALK ABOUT : By \textit{solving} And-or tree for Inhabitation tree
               we obtain that typed tree notation for typed $\lambda$-term,
               which must be firstly INTRODUCED.   \\
\textbf{TODO}: TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}
\\\\

\subsection{Our approach to \textit{solving} Inhabitation machine}

\subsection{Inhabitation Machine}

\section{Roadmap}
	\section{Conversion to SKI combinators}
